
---
title: "The influence of El Niño Southern Oscillation on the population dynamics of oceanic manta rays in Pacific Mexico "
author: "Madalena Cabral"
date: '2022'
output:
  html_document:
    code_folding: hide
    fig_caption: yes
    force_captions: yes
    highlight: pygments
    number_sections: yes
    theme: cerulean
    toc: yes
    toc_depth: 3
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '3'
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
    toc: yes
    toc_depth: 3
---


Here we use the capture histories to apply models of abundance with 3 environmental variables such as MEI(Multivariate ENSO index), SST(Sea Surface Temperature) and Chl-a (Chlorophyla a concentration). 


First we load the data from capture histories "Madalena CH Table.RData" for 48 months 
And we load "EnvData.RData" with environmental data for all 3 variables.

```{r}
load("Madalena CH Table.RData")
load("EnvData.RData")

EnviroData <- EnvData[,3:5]
EnviroData$time <- 1:52


library(RMark)
library(readxl)
library(ggplot2)
library(ggpubr)
library(xlsx) 

windowsFonts(A = windowsFont("Arial"))



```

To format the data with no covariates
```{r}
dp=process.data(table,model="POPAN")
ddl=make.design.data(dp)
str(ddl)
```


We use POPAN modeling for Open populations:

```{r}
CapHist <- table

CapHist$ch <- substr(table$ch,1,48)
CapHist <- CapHist[which(CapHist$ch != paste0(c(rep(0,48)),collapse="") ),] # With only 48 months, some of the capture histories at all zeros (we need to remove those)


dp=process.data(CapHist,model="POPAN")
ddl=make.design.data(dp)
str(ddl)

```

To merge MEI, SST and Chla into ddl:


```{r}
ddl$p=merge_design.covariates(ddl$p,EnviroData)
ddl$Phi=merge_design.covariates(ddl$Phi,EnviroData)
ddl$pent=merge_design.covariates(ddl$pent,EnviroData)

```

Now we want to run all possible combinations of variables and select a best model using AIC

Simple model for initial values, to improve model convergence:
the fixed=list() in p is where you fix detection probability to 0 for unsampled months. So time=c(9,10,11,12,21,22,23,24,33,34,35,36,45,46,47,48) for the unsampled summer months in our Revilla capture history. They are July, August, September and October and last cut November). 

```{r}
meses0 <- c(1,9,10,11,12,13,21,22,23,24,25,33,34,35,36,37,45,46,47,48)
Meses0<-as.factor(meses0)
initial=mark(dp,ddl,model.parameters=list(p=list(formula=~1,fixed=list(time=Meses0,value=0)),Phi=list(formula=~1),pent=list(formula=~1), N=list(formula=~1)))
```

To run analysis with all combinations of covariates we create the formulas for all Phi, p, pent and N and with create.model.list("POPAN") we can automaticaly create all possible combinations and then run all models and return as a list with class marklist

```{r}
do_analysis=function()
{
  # create formulas for Phi (survival probability) 
  Phi.dot=list(formula=~1)
  Phi.time=list(formula=~time)
  Phi.enso=list(formula=~MEI)
  Phi.chla=list(formula=~Chla)
  Phi.sst=list(formula=~SST)
  
  
  #create formulas for p (capture probability)
  p.dot=list(formula=~1, fixed=list(time=Meses0,value=0))
  p.time=list(formula=~time, fixed=list(time=Meses0,value=0))
  p.enso=list(formula=~MEI, fixed=list(time=Meses0,value=0))
  p.chla=list(formula=~Chla, fixed=list(time=Meses0,value=0))
  p.sst=list(formula=~SST, fixed=list(time=Meses0,value=0))
 
  #create formulas for pent (entry propability)
  pent.dot=list(formula=~1)
  pent.time=list(formula=~time)
  pent.enso=list(formula=~MEI)
  pent.chla=list(formula=~Chla)
  pent.sst=list(formula=~SST)
  
  
  #create formula for N. (population size) We keep this one constant
  N=list(formula=~1)
  
    # create all combinations
  cml=create.model.list("POPAN")
  
  # run all models and return as a list with class marklist
  results=mark.wrapper(cml,data=dp,ddl=ddl,output=FALSE,silent=TRUE, initial=initial)
  return(results)
}
revilla.results=do_analysis()
revilla.results
```

After looking at the model table, we can re-run the top models as selected by AIC criteria to pull out all of the parameter estimates. 

```{r}
top.model = mark(dp,ddl,model.parameters=list(p=list(formula=~time,fixed=list(time=c(1,9,10,11,12,13,21,22,23,24,25,33,34,35,36,37,45,46,47,48),value=0)),Phi=list(formula=~MEI),pent=list(formula=~1), N=list(formula=~1)))

second.model= mark(dp,ddl,model.parameters=list(p=list(formula=~time,fixed=list(time=c(1,9,10,11,12,13,21,22,23,24,25,33,34,35,36,37,45,46,47,48),value=0)),Phi=list(formula=~MEI),pent=list(formula=~Chla), N=list(formula=~1)))

```

To view superpopulation size

```{r}

top.model$results$derived$`N Population Size`

second.model$results$derived$`N Population Size`
```

To write table with monthly abundances

```{r}
write.table(top.model$results$derived$`N Population Size`,file="abundancetop.model.txt")
write.table(second.model$results$derived$`N Population Size`,file="abundancesecond.model.txt")
```
 
 
```{r}
summary(top.model,se=T)$reals$N
summary(second.model,se=T)$reals$N
```
 
For a table with all Parameter estimates

```{r}
parameter.estimates<-top.model$results$real
parameter.estimates

write.csv(parameter.estimates,"Param_Est_Real.csv",row.names = T)

parameter.estimates2<-second.model$results$real
parameter.estimates2

write.csv(parameter.estimates2,"Param_Est_Real.csv",row.names = T)

write.table(CapHist,file="abundancetop.model.txt")

```

```{r}
parameter.estimates<-top.model$results$real
parameter.estimates

parameter.estimates2<-second.model$results$real
parameter.estimates2
```


